---
title: "Coursera"
author: Val
date: March 22, 2017
output:
  md_document:
    variant: markdown_github
---
 
## Coursera, reproducible research, week 2

## Step 1: Read csv and convert to convenient format

```{r step0, echo=FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
```

```{r step1, echo=TRUE}
mdata <- read.csv("./data/activity.csv", stringsAsFactors = FALSE, header = TRUE) 
mdata$steps <- as.numeric(mdata$steps)
mdata$date <- as.Date(mdata$date, "%Y-%m-%d")

# group data by days
mdata.group_by_date <- mdata%>%
  group_by(date)%>% 
  summarise(steps = sum(steps, na.rm = TRUE))
```


## Step 2: histogram of the total number of steps per day
```{r step2, echo=TRUE}
p <- ggplot(mdata.group_by_date, aes(date, steps)) + geom_histogram(stat = "identity", na.rm = TRUE) + 
      geom_hline(aes(yintercept = mean(mdata.group_by_date$steps), colour = "red")) + 
      guides(color = guide_legend(title = "mean"))
plot(p)
```

## Step 3: Mean and median number of steps
```{r step3, echo=TRUE}
mmean <- mean(mdata.group_by_date$steps, na.rm = TRUE)
median <- median(mdata.group_by_date$steps, na.rm = TRUE)
mmean
median
```


## Step 4: Time series plot of the average number of steps taken
```{r step4, echo=TRUE}
mdata.group_by_interval <- mdata%>%
  group_by(interval)%>%
  summarise(steps = mean(steps, na.rm = TRUE))

p <- ggplot(mdata.group_by_interval, 
            aes(interval, steps), 
            type = "l") +  geom_line(aes(y = steps)) 
plot(p)
```

## Step 5: The 5-minute interval that, on average, contains the maximum number of steps
```{r step5, echo=TRUE}
max.int <- which(mdata.group_by_interval$steps == max(mdata.group_by_interval$steps))
mdata.group_by_interval[max.int,]
```
We could see that interval num 835 has max steps (~206)

## Step 6: Code to describe and show a strategy for imputing missing data
Missing values are replaced by average values from intervals

```{r step6, echo=TRUE}
count.na <- sum(is.na(mdata$steps))

# New dataset, without NA - mean by intervals
newdata <- merge(mdata, mdata.group_by_interval, by = "interval")
nas <- which(is.na(newdata$steps.x))
newdata[nas, "steps.x"] <- newdata[nas, "steps.y"]
colnames(newdata)[2] <- "steps"
# Remove unnessesary column
newdata <- newdata[, !names(newdata) == "steps.y"]
newdata <- newdata[with(newdata, order(date, interval)), ]
```

## Step 7: Histogram of the total number of steps taken each day after missing values are imputed

```{r step7, echo=TRUE}
newdata.group_by_date <- newdata%>%
  group_by(date)%>% 
  summarise(steps = sum(steps))

p <- ggplot(newdata.group_by_date, aes(date, steps)) + geom_histogram(stat = "identity", na.rm = TRUE) + 
  geom_hline(aes(yintercept = mean(newdata.group_by_date$steps), colour = "red")) + 
  guides(color = guide_legend(title = "mean"))
plot(p)
```

## Step 8: Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r step8, echo=TRUE}
Sys.setlocale("LC_TIME", "English")
newdata$weekdesc <- as.factor(ifelse(weekdays(newdata[,"date"]) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
table(newdata$weekdesc)
newdata.group_by_interval <- newdata%>%
  group_by(interval, weekdesc)%>%
  summarise(steps = mean(steps))

p <- ggplot(newdata.group_by_interval, 
    aes(interval, steps, colour = weekdesc), type = "l") + 
    geom_line(aes(y = steps)) + 
    facet_grid(weekdesc ~ .)
plot(p)
```
